Evaluation of pilot jobs for Apache Spark\\ applications on HPC clusters

### 2019-05-20 paper.tex

    % maybe remove
    \subsection{Pipeline robustness}
        While nodes failures can occur for various reasons, underestimation of 
        walltime may lead to early node termination in pilot scheduling systems.
        Therefore, it is crucial for the overlay cluster within the pilot system
        to be robust to node failures. In this series of experiments, we investigate
        Spark's fault-tolerance with respect to worker, master and driver termination.

        In our configuration, the driver is running in cluster deploy-mode, resulting
        in it being executed directly on a worker. This feature alone should not make
        it fault-tolerant to failures. However, using this feature coupled with the 
        \texttt{--supervise} flag, will.

@@

Although queuing time is largely responsible for makespan variations, it is not
always entirely responsible for the difference. Sometimes there are errors related
to workers not registering properly with the master. These occurrences lead to the 
application being processed entirely with a smaller number of available workers. As the task
delay added considers the maximum amount of parallel workers, functioning with less total workers
will significantly affect makespan. However, as seen in Figure~\ref{fig:mwall}
pilots are slower even with the same amount of average workers. A potential reasoning for this
may be the worker registration delay and the time required to transfer data
over after the application has already started. In batch scheduling, all workers are started in
parallel and the data is transferred automatically at the beginning of the pipeline as the maximum amount
of parallelism is available at driver start time. Pilots, on the other hand, may start workers in parallel,
but not necessarily all will be started at the same time. This means that the entire 
workflow may suffer the impacts of starting workers in sequence \TG{isn't it queuing time again?}. Moreover, there would also
be a delay with respect to transferring data from workers to newly added workers. Therefore,
it is only natural that pilots would have a bit more overhead than batch given the same
queuing time should pilots not all commence at once \TG{ok, but isn't the total amount of transfered data the same in both conditions?}.
 Furthermore, the pilot scripts also
have some startup overhead as all pilots start a master, a worker (registered to the main
master) and attempt to start a driver. Batch needs only start a single master and attempts to start
a driver only once.

The motivation to use pilots, however, is dynamic scheduling on HPC clusters. While we
investigated the queuing time differences with respect to differences in available resources,
the walltimes were kept static. Had walltime been underestimated to the point where the Slurm jobs
would terminate prior to application completion, pilots might have been preferable to batch
as batch scheduling has no mechanism to restart. Even if such a mechanism was in place, the entire
cluster would be shutdown and need to restart, which has additional overheads. With pilots, as long as 
there was at least one pilot alive, it would be possible to maintain the cluster and add additional pilots.
Spark provides built-it fault-tolerance for not only the workers, but the master and driver as well.

\tristan{Comment on the difficulties introduce by adding another layer in an already complex system:
HPC + Spark. It complexifies debugging.}

@@

\section{Conclusion}\label{sec:conclusion}

>>>>

    % maybe remove
    %\subsection{Pipeline robustness}
    %    While nodes failures can occur for various reasons, underestimation of 
    %    walltime may lead to early node termination in pilot scheduling systems.
    %    Therefore, it is crucial for the overlay cluster within the pilot system
    %    to be robust to node failures. In this series of experiments, we investigate
    %    Spark's fault-tolerance with respect to worker, master and driver termination.

    %   In our configuration, the driver is running in cluster deploy-mode, resulting
    %    in it being executed directly on a worker. This feature alone should not make
    %    it fault-tolerant to failures. However, using this feature coupled with the 
    %    \texttt{--supervise} flag, will.

@@

Although queuing time is largely responsible for makespan variations, it is not
always entirely responsible for the difference. Sometimes there are errors related
to executors not properly starting. These occurrences lead to the 
application being processed entirely with a smaller number of available executors. As the task
delay added considers the maximum amount of parallel workers, functioning with less total workers
will significantly affect makespan. However, as seen in Figure~\ref{fig:mwall}
pilots are slower even with the same amount of average workers. A potential reasoning for this
may be the worker registration delay and the time required to transfer data
over after the application has already started. In batch scheduling, all workers are started in
parallel and the data is transferred automatically at the beginning of the pipeline as the maximum amount
of parallelism is available at driver start time. Pilots, on the other hand, may start workers in parallel,
but not necessarily all will be started at the same time (i.e. delays in starting the workers once resources
have been allocated). This means that the entire 
workflow may suffer the impacts of starting workers in sequence rather than in parallel. Moreover, there would also
be a delay with respect to transferring data from workers to newly added workers. Therefore,
it is only natural that pilots would have a bit more overhead than batch given the same
queuing time should pilots not all commence at once, despite transferring the same amount of data.
 Furthermore, the pilot scripts also
have some startup overhead as all pilots start a master, a worker (registered to the main
master) and attempt to start a driver. Batch needs only start a single master and attempts to start
a driver only once.

The motivation to use pilots, however, is dynamic scheduling on HPC clusters. While we
investigated the queuing time differences with respect to differences in available resources,
the walltimes were kept static. Had walltime been underestimated to the point where the Slurm jobs
would terminate prior to application completion, pilots might have been preferable to batch
as batch scheduling has no mechanism to restart. Even if such a mechanism was in place, the entire
cluster would be shutdown and need to restart, which has additional overheads. With pilots, as long as 
there was at least one pilot alive, it would be possible to maintain the cluster and add additional pilots.
Spark provides built-it fault-tolerance for not only the workers, but the master and driver as well. However,
it is also likely that with faster queuing times, all resources would be allocated at the same time. If walltime
were to be underestimated, in this case, checkpointing and restarting the entire application from the last
checkpoint would be the only option. This too has limitations as the last checkpoint may not be recent, requiring
necessitating significant recomputation. It
might even be desirable to have many small resource requests in such a scenario, as with an increase in pilots,
it becomes less likely for all of them to be scheduled at once, as can be seen with 16 pilots. This would allow
for more chances for the application to reach completiong without failing entirely.

Many failures can be found, particularly with regards to the pilots. Many of these have been
a result of Spark failing to start properly. The driver either would not return a submission ID after
being started, or would remain in \texttt{SUBMITTED} status until the walltime expiration, having 
not processed any data. As these error become fewer with increased resources, it is possible they are
entirely related to lack of resources. Furthermore, it may also explain why batch submission always passes.
Not only are all the workers that were requested available when the pilot begins running, batch has one extra
executor available thanks to the driver process running directly on the process which had created it, rather than
a worker process. Conversely, it is not only the pilot applications that fail, the batch applications
do experience some failures as well. It can be seen in Figure~\ref{fig:makespansbeluga} that the batch application failed
twice: once in Configuration~2 and the other in Configuration~4. In both cases, the reason for failure was the same: lack of
available resources despite all executors having been assigned. These failures might be due to running on a cluster. Either
the resources were down, or other users might have been using them, resulting in the resources being unavailable to our application.

Running a Spark cluster atop of a Slurm allocation complexifies debugging, particularly if the application runs in cluster mode.
Spark worker logs, unless saved to network storage, are unavailable. Saving them to network storage also effectively slows down
the application as it is a slower storage than RAM and local disk. As a result, the Spark worker UI is also inaccessible. When running
in client mode, the driver output ends up it standard output/error. However, in cluster mode, the driver output is found stored within
the worker logs, which means the driver logs are also inaccessible without worker logs being written to network storage. Some HPC clusters
may have a file limit for user, meaning that the user must be sure that these logs coupled without other stored data does not exceed the file
limit.

@@

\section{Conclusion}\label{sec:conclusion}
Pilots do not provide significant advantage over batch systems using the same amount of resources. This is affected
by the overhead of dynamically adjusting the cluster size and the fact that queuing time of the batch applications took
as long as the pilots. These results, however, are to an extent cluster-dependent. It is expected that as the number
of resources requested increases, the speedup provided by pilots will increase. Furthermore, users executing on a smaller cluster 
may experience greater speedups using the same resource configuration.

Pilots are meant to allow for underestimation of resources, particularly walltime, which would not
be known to the user prior to execution. In its current state, \texttt{SPA} is not able to 
evaluate that functionality. Further experiments on master fault-tolerance, driver recovery, and checkpointing
will need to be conducted to determine if pilot jobs are even a viable solution for Spark applications running on 
HPC.

Currently, as client mode is not available in Spark Standalone clusters for PySpark applications,
using pilots on HPC would be limited to non-Python applications until there is a solution for 
Standalone mode.
